### 2008-08-05 doc/backend_interface/Backends-arc1.tex
Renamed files from Backends.* to Backends-arc1.* to be able to distinguish the back-ends' documentation of ARC0 from ARC1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {T.Frågåt},
  pdftitle ={Arc batch system back-end interface guide with support for GLUE 2},
  pdfsubject = {Description and developer's guide for ARC1},
  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Backend,LRMS,Infosystem,GLUE2},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

@@

{\raggedleft NORDUGRID-TECH-XX\par}

@@

\begin{center}
\textsc{\Large Arc batch system back-end interface guide with support for GLUE 2}
\par\end{center}{\Large \par}

@@

{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no} \par}

@@

This document describes the Advanced Resource Connector~\cite{arc1}
batch system back-end infrastructure. It aims to describe the infrastructure
in enough detail that a developer can add a new local resource management
system (LRMS) to the ARC middleware. It also intends to serve as a
reference manual describing existing batch system interfaces. Note
that the current solution described in this manual is not a real interface,
it is too diffuse to be that.

The batch system back-ends are what tie the ARC grid middleware (through
(ARC Resource-coupled EXecution service, A-REX~\cite{arex}) to the underlying
cluster management system or LRMS. The back-ends consist of set of shell and
Perl scripts whose role are twofold:

\begin{enumerate}
\item to allow the GM, which is resided in A-REX, to manipulate jobs on the LRMS including job submit,
cancel operations etc.
\item to collect information about jobs, users, batch system and the cluster
itself for the Information System. 
\end{enumerate}
The former will be referred to as the job control back-end interface
while the latter is the information system interface of the batch
system back-ends. These two will be treated separately in the following
sections.

@@

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
within the NorduGrid subversion tree~\cite{svn}. 
The job control interface scripts can be found under \texttt{arc1/trunk/src/services/a-rex/lrms/}
while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.

The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.

@@

Job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
It takes care of preparing a native batch system submission script,
managing the actual submission of the batch system job, cancellation
of job on request and scanning for completed batch jobs. Besides the
LRMS job control interface it is also the GM which provides e.g. the
data staging and communication with the grid client, provides RTE
environments, arranges file staging (to the node via LRMS capability),
dealing with stdout/stderr, etc. The job control batch system interface
of the GM requires three programs. These programs can be implemented
any way the designer sees fit, but all the existing back-end interfaces
use shell scripting for portability and ease of tailoring to a specific
site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
and scan-LRMS-job. LRMS is replaced with the short hand name for the
LRMS e.g. cancel-pbs-job. The scripts are described one by one in
the following subsections. Useful information can also be found in
the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
Runtime Environment\char`\"{} of the Grid-Manager guide~\cite{gm}.

@@

The submit program is the most involved. It is called by the GM once
a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
it is given the grami file as argument on execution. Submit-LRMS-job
then has to set up the session directories, run-time environment and
anything else needed. Then it should submit the job to the local LRMS.
This is normally done by generating a native job script for the LRMS
and then running the local submit command, but it could also be done
through an API if the LRMS supports it.

@@

Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
jobs that have finished. Once it has found a finished job it should
write the exit-code of that job to the file job.\{gridid\}.lrms\_done
in the ARC jobstatus directory. Then it should call the gm-kick program
the GM will then notice that the job has finished and start finalizing
the job.

Generally two approaches are taken to find jobs that are done. One
is asking the LRMS. First all started grid jobs are found in the jobstatus
directory%
\footnote{normally /var/spool/nordugrid/jobstatus/, but can be set via the controldir
variable of arc.conf%
}. This can be done by checking if the status is INLRMS in the jobs
status file%
\footnote{job.\{gridid\}.status%
}. Then the LRMS is asked for the status of those jobs. If they are
done they are marked as such and the GM is activated. The problem
with this approach is that for most LRMSs the information about finished
jobs are only available for a short period after the job finished.
Therefore appropriate steps have to be taken if the job is known to
have been started but is no longer present in the LRMS. The normal
approach is to analyze the jobs status output in the session directory.

The second approach is to parse the LRMSs log files. This method has
some drawbacks e.g. the GM has to be allowed read access to the logs.
The back-end will then have to remember where in the log it was last
time it ran. This information will have to be stored in a file some
where on the front-end.

@@

\begin{quote}
Parses a config file. It returns exit status 1 if the file cannot
be read, or 0 otherwise. Badly formed lines are silently ignored.
Option values can be surrounded by single quotes or double quotes.
Values without quotes are also accepted. Multi-valued options are
not supported currently. Only the last defined value is retained. 
\end{quote}
\texttt{config\_import\_section <section\_name>}

@@

The main purpose of the information system batch interface is to populate
the nordugrid information model and the GLUE2 model with locally collected
information obtained from the batch system. Important to recall that the
locally collected information is broader than what the batch-system can
offer, information taken from the grid-layer (mostly grid manager), from
the front-end machine and from the \texttt{arc.conf} configuration
file are also needed and used to populate the nordugrid information
model~\cite{is} and the GLUE2 schema~\cite{glue2}.

The information system interface consists of a set of Perl scripts which
generates an XML output to \texttt{STDOUT}.

The following subsection describes the LRMS interface part of the
information system. To support a particular LRMS a dedicated Perl module
should be written that implements the interface presented in the next
subsections. Furthermore hooks should be added to the \texttt{LRMS.pm} module,
so that it uses the correct Perl module depending on the LRMS type.

The information collector scripts are divided between several separate scripts:
\begin{itemize}
\item \texttt{cluster+qju.pl} - driver for information collection. It calls all other information collectors and prints results in XML. The information collection is done by one single invocation of this script
\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
\item \texttt{InfoChecker.pm} - used by \texttt{InfoCollector} to validate options and results against a ``schema''.
\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...) by calling the appropriate LRMS plugin. It also does validation of input options to the plugin and of the data returned by the plugin. 
\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
\end{itemize}

All information about the hosting environment and queues is collected by one invocation of the Perl script named
\texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes
two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.

@@

The \texttt{lrms\_info} function should then return a
hash of hashes with the sctructure described in table \ref{t:lrms_info}.

@@

\begin{table}
\caption{The \texttt{users\_info} hash}
\label{t:users_info}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
freecpus & number of freely available CPUs with their time limits in minutes
(see ComputingService.FreeSlotsWithDuration in\cite{glue2})\\
\hline
queuelength & estimated queue length for the specified user\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

This section presents the batch system specific implementation details
including information on supported versions, constraints on batch
system configuration, known limitations, \texttt{arc.conf} parameters
and list of batch system features being utilized within the interface
are described.

@@

PBS is a very powerful Local Resource Manager System with dozens of
configurable options. Server, queue and node attributes can be used
to configure the cluster's behaviour. In order to correctly interface
PBS to ARC (mainly the information provider scripts) there are a couple
of configuration REQUIREMENTS asked to be implemented by the local
system administrator:

\begin{enumerate}
\item The computing nodes MUST be declared as cluster nodes (job-exclusive),
at the moment time-shared nodes are not supported by the ARC setup.
If you intend to run more than one job on a single processor then
you can use the virtual processor feature of PBS. 
\item For each queue, you MUST set one of the max\_user\_run or max\_running
attributes and its value SHOULD BE IN AGREEMENT with the number of
available resources (i.e. don't set the max\_running = 10 if you have
only six (virtual) processors in your system). If you set both max\_running
and max\_user\_run then obviously max\_user\_run has to be less equal
than max\_running. 
\item For the time being, do NOT set server limits like max\_running, please
use queue-based limits instead. 
\item Avoid using the max\_load and the ideal\_load directives. The node's
mom config file (<PBS home on the node>/mom\_priv/config) should not
contain any max\_load or ideal\_load directives. PBS closes down a
node (no jobs are allocated to it) when the load on the node reaches
the max\_load value. The max\_load value is meant for controlling
time-shared nodes. In case of job-exclusive nodes there is no need
for setting these directives, moreover incorrectly set values can
close down your node. 
\item Routing queues are not supported, those can't be used within ARC.
\end{enumerate}
Additional useful configuration hints:

@@

\begin{itemize}
\item The PBS batch system back-end is enabled via setting the \texttt{lrms=\char`\"{}pbs\char`\"{}}
in the \texttt{{[}common]} configuration block. No need to specify
the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
keyword as lrms configuration value.
\item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path to the qstat,pbsnodes,qmgr etc PBS
binaries.
\item \texttt{pbs\_log\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path of the PBS server logfiles which are
used by the Grid-Manager to determine whether a PBS job is completed.
If not specified, the Grid-Manager will use the \texttt{qstat} command
to find completed jobs. 
\item \texttt{lrmsconfig} from the \texttt{{[}cluster]} block can be used
as an optional free text field to describe further details about the
PBS configuration (e.g. \texttt{lrmsconfig=\char`\"{}single job per
processor\char`\"{}}).
\item \texttt{dedicated\_node\_string} from the \texttt{{[}cluster]} block
specifies the string which is used in the PBS node config to distinguish
the grid nodes from the rest. Suppose only a subset of nodes are available
for grid jobs, and these nodes have a common \texttt{node property}
string, this case the \texttt{dedicated\_node\_string} should be set
to this value and only the nodes with the corresponding PBS \texttt{node
property} are counted as grid enabled nodes. Setting the \texttt{dedicated\_node\_string}
to the value of the PBS \texttt{node property} of the grid-enabled
nodes will influence how the totalcpus, user freecpus is calculated.
No need to set this attribute if the cluster is fully available for
the grid and the PBS configuration does not use the \texttt{node property}
method to assign certain nodes to grid queues.
\item \texttt{{[}queue/queuename]} block. For each grid-enabled (or grid
visible) PBS queue a corresponding \texttt{{[}queue]} block must be
defined. \texttt{queuename} should be the PBS queue name.
\item \texttt{scheduling\_policy} from the \texttt{{[}queue/queuename]}
block describes the scheduling policy of the queue. PBS by default
offers the FIFO scheduler, many sites run the MAUI. At the moment
\texttt{FIFO} \& \texttt{MAUI} are supported values. If you have a
MAUI scheduler you should specify the \char`\"{}MAUI\char`\"{} value
since it modifies the way the queue resources are calculated. By default
the \char`\"{}FIFO\char`\"{} scheduler type is assumed.
\item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
sets the path of the maui commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
is specified as \texttt{scheduling\_policy} value. This parameter
can be set in the \texttt{{[}common]} block as well. 
\item \texttt{queue\_node\_string} of the \texttt{{[}queue/queuename]} block
can be used similar to the \texttt{dedicated\_node\_string}. In PBS
you can assign nodes to a queue (or a queue to nodes) by using the
\texttt{node property} PBS node configuration method and assigning
the marked nodes to the queue (setting the \texttt{resources\_default.neednodes}
= \texttt{queue\_node\_string} for that queue). This parameter should
contain the \texttt{node property} string of the queue-assigned nodes.
Setting the \texttt{queue\_node\_string} changes how the queue-totalcpus,
user freecpus are determined for this queue. 
\end{itemize}

@@

The job control batch interface makes use of the \texttt{qsub} command
to submit native PBS jobscripts to the batch system. The following
options are used:

@@

The information system interface utilizes the \texttt{qstat -f -Q
queuename} and \texttt{qstat -f queuename} commands to obtain detailed
job and queue information. \texttt{qmgr -c \char`\"{}list server\char`\"{}}
is used to determine PBS flavour and version. The \texttt{pbsnodes}
command is used to calculate total/used/free cpus within the cluster.
In case of a Maui scheduller the \texttt{showbf} command is used to
determine user freecpu values. All these external PBS commands are
interfaced via parsing the commands' output.

@@

The Fork back-end is a simple back-end that interfaces to the local
machine i.e. there is no batch system underneath. It simply forks
the job, hence the name. The back-end then uses standard posix commands(e.g.
ps or kill) to manage the job.

@@

Since fork is a simple back-end and does not use any batch system,
there is no specific configuration needed of the underlying system.

@@

\begin{itemize}
\item The Fork back-end is enabled by setting \texttt{lrms=\char`\"{}fork\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item The queue must be named {}``fork'' in the queue section.
\item fork\_job\_limit=\char`\"{}cpunumber\char`\"{}, this option is used
to set the number of running grid jobs on the fork machine, allowing
a multi core machine to use some or all of its cores for Grid jobs.
The default value is 1. 
\end{itemize}

@@

The Fork backend implements an interface to the {}``fork'' unix
command which is not a batch system. Therefore the backend should
rather be seen as an interface to the operating system itself. Most
of the {}``batch system values'' are determined from the operating
system (e.g. cpu load) or manually set in the configuration file.

Fork is not a batch system, therefore many of the queue specific attributes
or detailed job information is not available. The support for the
{}``Fork batch system'' was introduced so that quick deployments
and testing of the middleware can be possible without dealing with
deployment of a real batch system since fork is available on every
unix box. The Fork backend is not recommended to be used in production.
The backend by its nature, has lots of limitations, for example does
not support parallel jobs.

@@

Sun Grid Engine, formerly known as Codine, is an open source batch
system maintained by Sun. Runs on Linux, Solaris.

@@

The SGE backend's commands are similar to the PBS commands. These
commands are used in the code:

@@

\begin{itemize}
\item qconf -spl (show a list of all currently defined parallel environments)
\item qconf -sql (show a list of all queues)
\item qconf -sep (show a list of all licensed processors/slots)
\item qstat -g c (display cluster queue summary)
\item qconf -sconf global (show global configuration)
\item qconf -sq queue\_name (show the given queue cofiguration)
\end{itemize}
Other:

@@

Multi-CPU support is not well tested. All users are shown with the
same quotas in infosys, even if they are mapped to different local
users. The requirement that one ARC queue maps to one SGE queue is
too restrictive, as the SGE's notion of a queue differs widely from
ARC's definition. The flexibility available in SGE for defining policies
is difficult to accurately translate into nordugrid's information
schema. The closest equivalent of nordugrid-queue-maxqueuable is a
per-cluster limit in SGE, and the value of nordugrid-queue-localqueued
is not well defined if pending jobs can have multiple destination
queues.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {T.Frågåt},
  pdftitle ={Arc batch system back-end interface guide with support for GLUE 2},
  pdfsubject = {Description and developer's guide for ARC1},
  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Back-end,LRMS,Infosystem,GLUE2},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

@@

{\raggedleft NORDUGRID-TECH-18\par}

@@

\begin{center}
\textsc{\Large Arc batch system back-end interface guide with support for Glue 2}
\par\end{center}{\Large \par}

@@

{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomas.fragat@fys.uio.no} \par}

@@

This document describes the next generation Advanced Resource Connector~\cite{arc1} (ARC)
batch system back-end infrastructure. It aims to describe the infrastructure
in enough detail that a developer can add a new local resource management
system (LRMS) to the ARC middleware. It also intends to serve as a
reference manual describing existing batch system interfaces. Note
that the current solution described in this manual is not a real interface,
it is too diffuse to be that.

The batch system back-ends are what tie the ARC grid middleware (through the
ARC Resource-coupled EXecution service, A-REX~\cite{arex}) to the underlying
cluster management system or LRMS. The back-ends consist of set of a shell and
Perl scripts whose role are twofold:

\begin{enumerate}
\item to allow the GridManager (GM), which is resided in A-REX, to manipulate
jobs in the LRMS including job submit, cancel operations etc.
\item to collect information about jobs, users, the batch system and the cluster
itself for the Information System. 
\end{enumerate}
The former will be referred to as the job control back-end interface
while the latter is the information system interface of the batch
system back-ends. These two will be treated separately in the following
sections.

@@

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
within the NorduGrid subversion tree~\cite{svn}. 
The job control interface scripts can be found under 
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/lrms/}
\end{quote}
while the information collectors are located in
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
\end{quote}
The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.

@@

The job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
It takes care of preparing a native batch system submission script,
managing the actual submission of the batch system job, cancellation
of job on request and scanning for completed batch jobs. Besides the
LRMS job control interface it is also the GM which provides e.g. the
data staging and communication with the grid client, provides RTE
environments, arranges file staging (to the node via LRMS capability),
dealing with stdout/stderr, etc. The job control batch system interface
of the GM requires three programs. These programs can be implemented
any way the designer sees it fits, but all the existing back-end interfaces
use shell scripting for portability and ease of tailoring to a specific
site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
and scan-LRMS-job where LRMS is replaced with the short hand name for the
LRMS; e.g. cancel-pbs-job. The scripts are described one by one in
the following subsections. Useful information can also be found in
the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
Runtime Environment\char`\"{} of the Grid-Manager guide~\cite{gm}.

@@

The submit program is the most involved. It is called by the GM once
a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
it is given the GRAMi file as argument on execution. Submit-LRMS-job
then has to set up the session directories, run-time environment and
anything else needed. Then it submits the job to the local LRMS.
This is normally done by generating a native job script for the LRMS
and then running the local submit command, but it can also be done
through an API if the LRMS supports it.

@@

Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
jobs that have finished. Once it has found a finished job it will
write the exit-code of that job to the file job.\{gridid\}.lrms\_done
in the ARC jobstatus directory. Then it will call the gm-kick program
to notify GM about the finished job. Subsequently, the GM starts finalizing
the job.

Generally, two approaches are taken to find jobs that are done. One
is to ask the LRMS. First all started grid jobs are found in the jobstatus
directory%
\footnote{normally /var/spool/nordugrid/jobstatus/, but can be set via the controldir
variable of arc.conf%
}. This can be done by checking if the status is "INLRMS" in the jobs
status file%
\footnote{job.\{gridid\}.status%
}. Then the LRMS is asked for the status of those jobs. If they are
done they are marked as such and the GM is activated. The problem
with this approach is that for most LRMSs the information about finished
jobs are only available for a short period after the job finished.
Therefore appropriate steps have to be taken if the job is known to
have been started but is no longer present in the LRMS. The normal
approach is to analyze the jobs status output in the session directory.

The second approach is to parse the LRMSs log files. This method has
some drawbacks, e.g. the GM has to be allowed read access to the logs.
The back-end will then have to remember where in the log it was last
time it ran. This information will have to be stored in a file somewhere
on the front-end.

@@

\begin{quote}
Parses a config file. It returns exit status 1 if the file cannot
be read, or 0 otherwise. Badly formed lines are silently ignored.
Option values can be surrounded by single quotes or double quotes.
Values without quotes are also accepted. Currently, multi-valued options are
not supported. Only the last defined value is retained. 
\end{quote}
\texttt{config\_import\_section <section\_name>}

@@

The main purpose of the information system batch interface is to populate
the NorduGrid information model and the GLUE2 model with locally collected
information obtained from the batch system. It is important to recall that the
locally collected information is broader than what the batch-system can
offer, information taken from the grid-layer (mostly A-REX), from
the front-end machine, and from the \texttt{arc.conf} configuration
file are also needed and used to populate the NorduGrid information
model~\cite{is} and the GLUE2 schema~\cite{glue2}.

The information system interface consists of a set of Perl scripts which
generates an XML output to \texttt{STDOUT} by one invocation of the Perl script named
\texttt{cluster+qju.pl}. The XML output includes two representations of the output data,
namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.

The information collector scripts are divided between several separate scripts:
\begin{itemize}
\item \texttt{cluster+qju.pl} - driver for information collection. It calls all other information collectors and prints the results in XML. The information collection is done by one single invocation of this script.
\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
\item \texttt{InfoChecker.pm} - used by \texttt{InfoCollector} to validate options and results against a ``schema''.
\item \texttt{GMJobsInfo.pm} - collects information about jobs from GM status files.
\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments, etc.)
\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits, etc.) by calling the appropriate LRMS plugin. It also does validation of input options to the plugin and of the data returned by the plugin. 
\item \texttt{<BATCH\_SYSTEM\_NAME>.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework.
\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NorduGrid schema.
\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema.
\end{itemize}

The following subsection describes the LRMS interface part of the
information system. To support a particular LRMS a dedicated Perl module
should be written that implements the interface presented in the next
subsections. Furthermore, hooks should be added to the \texttt{<BATCH\_SYSTEM\_NAME>.pm} module,
so that it uses the correct Perl module depending on the LRMS type.

@@

The \texttt{lrms\_info} function should then return a
hash of hashes with the structure described in table \ref{t:lrms_info}.

@@

\begin{table}
\caption{The \texttt{users\_info} hash}
\label{t:users_info}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
freecpus & number of freely available CPUs with their time limits in minutes
(see ComputingService.FreeSlotsWithDuration in~\cite{glue2})\\
\hline
queuelength & estimated queue length for the specified user\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

This section presents the batch system specific implementation details
including information on supported versions, constraints on batch
system configuration, known limitations, \texttt{arc.conf} parameters,
and a list of batch system features being utilized within the interface
are described.

@@

PBS is a very powerful LRMS with dozens of
configurable options. Server, queue and node attributes can be used
to configure the cluster's behaviour. In order to correctly interface
PBS to ARC (mainly the information provider scripts) there are a couple
of configuration REQUIREMENTS asked to be implemented by the local
system administrator:

\begin{enumerate}
\item The computing nodes MUST be declared as cluster nodes (job-exclusive),
at the moment time-shared nodes are not supported by the ARC setup.
If you intend to run more than one job on a single processor then
you can use the virtual processor feature of PBS. 
\item For each queue, you MUST set one of the max\_user\_run or max\_running
attributes and its value SHOULD BE IN AGREEMENT with the number of
available resources (i.e. don't set the max\_running = 10 if you have
only six (virtual) processors in your system). If you set both max\_running
and max\_user\_run then obviously max\_user\_run has to be less or equal
to max\_running.
\item For the time being, do NOT set server limits like max\_running, please
use queue-based limits instead. 
\item Avoid using the max\_load and the ideal\_load directives. The node's
mom configuration file (<PBS home on the node>/mom\_priv/config) should not
contain any max\_load or ideal\_load directives. PBS closes down a
node (no jobs are allocated to it) when the load on the node reaches
the max\_load value. The max\_load value is meant for controlling
time-shared nodes. In case of job-exclusive nodes there is no need
for setting these directives, moreover incorrectly set values can
close down your node. 
\item Routing queues are not supported, those cannot be used within ARC.
\end{enumerate}
Additional useful configuration hints:

@@

\begin{itemize}
\item The PBS batch system back-end is enabled via setting the \texttt{lrms=\char`\"{}pbs\char`\"{}}
in the \texttt{{[}common]} configuration block. No need to specify
the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
keyword as LRMS configuration value.
\item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path to the qstat,pbsnodes,qmgr etc PBS
binaries.
\item \texttt{pbs\_log\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path of the PBS server logfiles which are
used by the GM to determine whether a PBS job is completed.
If not specified, the GM will use the \texttt{qstat} command
to find completed jobs. 
\item \texttt{lrmsconfig} from the \texttt{{[}cluster]} block can be used
as an optional free text field to describe further details about the
PBS configuration (e.g. \texttt{lrmsconfig=\char`\"{}single job per
processor\char`\"{}}).
\item \texttt{dedicated\_node\_string} from the \texttt{{[}cluster]} block
specifies the string which is used in the PBS node config to distinguish
the grid nodes from the rest. Suppose only a subset of nodes are available
for grid jobs, and these nodes have a common \texttt{node property}
string, this case the \texttt{dedicated\_node\_string} should be set
to this value and only the nodes with the corresponding PBS \texttt{node
property} are counted as grid enabled nodes. Setting the \texttt{dedicated\_node\_string}
to the value of the PBS \texttt{node property} of the grid-enabled
nodes will influence how the totalcpus, user freecpus is calculated.
No need to set this attribute if the cluster is fully available for
the grid and the PBS configuration does not use the \texttt{node property}
method to assign certain nodes to grid queues.
\item \texttt{{[}queue/queuename]} block. For each grid-enabled (or grid
visible) PBS queue a corresponding \texttt{{[}queue]} block must be
defined. \texttt{queuename} should be the PBS queue name.
\item \texttt{scheduling\_policy} from the \texttt{{[}queue/queuename]}
block describes the scheduling policy of the queue. PBS by default
offers the FIFO scheduler, many sites run the MAUI. At the moment
\texttt{FIFO} \& \texttt{MAUI} are supported values. If you have a
MAUI scheduler you should specify the \char`\"{}MAUI\char`\"{} value
since it modifies the way the queue resources are calculated. By default
the \char`\"{}FIFO\char`\"{} scheduler type is assumed.
\item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
sets the path of the maui commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
is specified as \texttt{scheduling\_policy} value. This parameter
can be set in the \texttt{{[}common]} block as well. 
\item \texttt{queue\_node\_string} of the \texttt{{[}queue/queuename]} block
can be used similar to the \texttt{dedicated\_node\_string}. In PBS
you can assign nodes to a queue (or a queue to nodes) by using the
\texttt{node property} PBS node configuration method and assigning
the marked nodes to the queue (setting the \texttt{resources\_default.neednodes}
= \texttt{queue\_node\_string} for that queue). This parameter should
contain the \texttt{node property} string of the queue-assigned nodes.
Setting the \texttt{queue\_node\_string} changes how the queue-totalcpus,
user freecpus are determined for this queue. 
\end{itemize}

@@

The job control batch interface makes use of the \texttt{qsub} command
to submit native PBS job scripts to the batch system. The following
options are used:

@@

The information system interface utilizes the \texttt{qstat -f -Q
queuename} and \texttt{qstat -f queuename} commands to obtain detailed
job and queue information. \texttt{qmgr -c \char`\"{}list server\char`\"{}}
is used to determine PBS flavour and version. The \texttt{pbsnodes}
command is used to calculate total/used/free cpus within the cluster.
In case of a MAUI scheduler the \texttt{showbf} command is used to
determine user freecpu values. All these external PBS commands are
interfaced via parsing the commands' output.

@@

The Fork back-end is a simple back-end that interfaces to the local
machine, i.e.: there is no batch system underneath. It simply forks
the job, hence the name. The back-end then uses standard posix commands
(e.g. ps or kill) to manage the job.

@@

Since fork is a simple back-end and does not use any batch system,
there is no specific configuration needed for the underlying system.

@@

\begin{itemize}
\item The Fork back-end is enabled by setting \texttt{lrms=\char`\"{}fork\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item The queue must be named {}``fork'' in the queue section.
\item fork\_job\_limit=\char`\"{}cpunumber\char`\"{}, this option is used
to set the number of running grid jobs on the fork machine, allowing
a multi-core machine to use some or all of its cores for Grid jobs.
The default value is 1. 
\end{itemize}

@@

The Fork back-end implements an interface to the {}``fork'' UNIX
command which is not a batch system. Therefore the back-end should
rather be seen as an interface to the operating system itself. Most
of the {}``batch system values'' are determined from the operating
system (e.g. cpu load) or manually set in the configuration file.

Since Fork is not a batch system, many of the queue specific attributes
or detailed job information is not available. The support for the
{}``Fork batch system'' was introduced so that quick deployments
and testing of the middleware can be possible without dealing with
deployment of a real batch system since fork is available on every
UNIX box. The "Fork back-end" is not recommended to be used in production.
The back-end by its nature, has lots of limitations, for example it does
not support parallel jobs.

@@

Sun Grid Engine, formerly known as Codine, is an open source batch
system maintained by Sun. It runs on Linux, and Solaris.

@@

The SGE back-end's commands are similar to the PBS commands. These
commands are used in the code:

@@

\begin{itemize}
\item qconf -spl (show a list of all currently defined parallel environments)
\item qconf -sql (show a list of all queues)
\item qconf -sep (show a list of all licensed processors/slots)
\item qstat -g c (display cluster queue summary)
\item qconf -sconf global (show global configuration)
\item qconf -sq queue\_name (show the given queue configuration)
\end{itemize}
Other:

@@

Multi-CPU support is not well tested. All users are shown with the
same quotas in the information system, even if they are mapped to different local
users. The requirement that one ARC queue maps to one SGE queue is
too restrictive, as the SGE's notion of a queue differs widely from
ARC's definition. The flexibility available in SGE for defining policies
is difficult to accurately translate into NorduGrid's information
schema. The closest equivalent of nordugrid-queue-maxqueuable is a
per-cluster limit in SGE, and the value of nordugrid-queue-localqueued
is not well defined if pending jobs can have multiple destination
queues.
